---
title: "MetabolicResults"
output: html_document
---

## clear workspace, import data including new weight data and merge dataframes
```{r}
rm(list=ls())


library(tidyverse)
library(lme4)
library(emmeans)

dat <- read.csv('C:/Users/Daniel.Feeney/Dropbox (Boa)/FBS Abstract/MetResults13.csv') # With 13 time points
weightNew <- read.csv('C:/Users/Daniel.Feeney/Dropbox (Boa)/FBS Abstract/Metabolic (31 of 31)/MetabolicTrialResults/weights2.csv') #correcting for different mass each day
weightNew <- rename(weightNew, Subject = ï..Subject)
weightNew <- rename(weightNew, Config = Condition)

dat <- merge(dat, weightNew)
dat$TimePoints <- as.factor(dat$TimePoints)
dat['EEmTmp'] <- dat$EEm * dat$Ratio

dat$TimePoints <- factor(dat$TimePoints, levels = c("TP1","TP2","TP3","TP4","TP5","TP6","TP7","TP8","TP9","TP10", "TP11","TP12","TP13"))

ggplot(data = dat, mapping = aes(x = as.factor(TimePoints), y = EEmTmp, fill = Config)) + geom_point(size = 2, aes(color = Config)) + facet_wrap( ~ Subject ) + ylab('Energetic Expenditure (W/kg)') + xlab('Time Point') + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```



### Use token outlier removal and remove non-physiological energetic expenditures (below 5). Multiply EEm (after removing outliers, EEm2 by the ratio of mass)
```{r}
### Outlier removal based on 25 quantile - 1.5* IQR or 75 + 1.5*IQR ###
remove_outliers <- function(x, na.rm = TRUE, ...) {
  mn <- mean(x, na.rm = TRUE)
  std <- sd(x, na.rm = TRUE)
  y <- x
  y[x < (mn - 1.96*std)] <- NA
  y[x > (mn + 1.96*std)] <- NA
  y
}
# Level
dat <- dat %>%
  group_by(Subject, Config) %>%
  mutate(
    EEm2 = remove_outliers(EEm)
  )
dat <- subset(dat, dat$EEm2 > 5) # removing one non physiological rates not detected by the outlier
dat['EEmFixed'] <- dat$EEm2 * dat$Ratio

ggplot(data = dat, mapping = aes(x = as.factor(TimePoints), y = EEmFixed, fill = Config)) + geom_point(aes(color = Config)) +
  facet_wrap( ~ Subject ) 
```

## Run simple model with EEm on configuration and time points with a sub-specific intercept and subject-specific slope between configs. Adding temperature into the model to control for any small changes we saw
```{r}
mod1 <- lmer(EEmFixed ~ Config + as.factor(TimePoints) + Temp + (1|Subject), data = dat)
summary(mod1)
modnull <- lmer(EEmFixed ~ + as.factor(TimePoints) + Temp + (1|Subject), data = dat)
anova(mod1,modnull)


conditions.emm <- emmeans(mod1, "Config")
contrast(conditions.emm, "trt.vs.ctrl", ref = "SL") #p-value of 0.036 for dual dial vs lace, estimated 1.6% improvement

```
```{r}
ggplot(data = dat, mapping = aes(x = as.factor(TimePoints), y = Temp, color = Config)) + geom_point() + facet_wrap(~Subject)

```


# Below is removing subjects based on temperature. I do not think we should do this because it lowers our statistical power.
# Subjects 26, 29, and 30 have high temperatures for the SL day (greater than 1 degree difference). Let's try removing them and don't include Temperature in the model. 
# Subject 4 has a high temp for DD. 
# The question is adding tempature into the model sufficient? If so, we don't need to drop these subjects. 
# Asking a physiologist, they said slight temperature fluctuations should NOT impact energetic expenditure that much
```{r}
#reducedDat <- subset(dat, !( dat$Subject == 26 & dat$Config == 'SL'))
reducedDat <- subset(dat, !( dat$Subject == 29 & dat$Config == 'SL'))
reducedDat <- subset(reducedDat, !( reducedDat$Subject == 30 & reducedDat$Config == 'SL'))
reducedDat <- subset(reducedDat, !( reducedDat$Subject == 4 & reducedDat$Config == 'DD'))


mod1 <- lmer(EEmFixed ~ Config + Temp + as.factor(TimePoints) + (1+Config|Subject), data = reducedDat)
summary(mod1)
modnull <- lmer(EEmFixed ~ Temp + as.factor(TimePoints)  + (1+Config|Subject), data = reducedDat)
anova(mod1,modnull) #not significant


conditions.emm <- emmeans(mod1, "Config")
contrast(conditions.emm, "trt.vs.ctrl", ref = "SL")


ggplot(data = reducedDat, mapping = aes(x = as.factor(TimePoints), y = EEmFixed, fill = Config)) + geom_point(aes(color = Config)) + facet_wrap( ~ Subject )
```

## We do not see sig differences when we drop those three subject time points


# Look at pre/post only
## Differences in met cost between early and later in testing. Found differences between last
## few minutes and first few minutes for 4 observations per person per shoe

```{r}
diffDat <- read.csv('C:/Users/Daniel.Feeney/Dropbox (Boa)/FBS Abstract/MetDiff.csv')
diffDat$Config <- as.factor(diffDat$Config)
diffDat$TimePoints <- as.factor(diffDat$TimePoints)
diffDat$Subject <- as.factor(diffDat$Subject)


diffDat <- subset(diffDat, !( diffDat$Subject == 28 &  diffDat$Config == 'DD'))
diffDat <- subset(diffDat, !( diffDat$Subject == 16 &  diffDat$Config == 'SL'))

ggplot(data = diffDat, mapping = aes(x = Subject, y = differences, color = Config)) + geom_boxplot() + geom_hline(yintercept = 0) 

mod1 <- lmer(differences ~ Config + (1|Subject), data = diffDat)
summary(mod1)
modnull <- lmer(differences ~ (1|Subject), data = diffDat)
anova(mod1,modnull) 

conditions.emm <- emmeans(mod1, "Config")
contrast(conditions.emm, "trt.vs.ctrl", ref = "SL") #p-value of 0.12 for DD to lace

```



